---
title: "Resample Tool With R"
output: html_notebook
---

# Reading in and preparing data 

When running the notebooks on VITO’s servers, the user can follow the directions shown in 
https://nbviewer.jupyter.org/github/VITObelgium/notebook-samples/raw/master/datasets/probav/Reading%20PROBA-V%20using%20R.ipynb to find locations of PROBA-V products. Doing so, the data set (netCDF files) will be located and available in the VITO's servers.

Alternatively, when working locally, the R-user can choose to automatically download Copernicus land data (https://land.copernicus.eu/global/) using the functions found in https://github.com/cgls/Copernicus-Global-Land-Service-Data-Download-with-R. 

Once the data set is available, *raster* package functionality is used to prepare it for resampling. 

```{r}
library(raster)

# ndvi_files is a list of the available files (netCFD)
# r300m <- substring(ndvi_files[1],6)  # ndvi300m is the filename of the object to be processed
r <- raster(r300m)

```


The original netCDF file might contain flagged data corresponding to NoData (NA), etc, which needs to be dealt with. In case of the NDVI products and netCDF files, assigned values > 250 are flagged and need to be converted into NA. When netCDF files are read in as a *Raster*\* object and values transformed into real NDVI values, the cuttoff for flagged cells is 0.92. Therefore, all cells > 0.92 have to be set to NA.

```{r}

if(grepl("ndvi", ndvi_files[1], ignore.case = TRUE)){
  cuttoff_flag <- 0.92
}else{
  stop("Please define 'cuttoff_flag' (cuttoff value for flagged cells) for the Raster* object")
}

r[r > cuttoff_flag] <- NA

```



# Resampling using the aggregation approach

There are several approaches to resample data to a coarser resolution. The area-based aggregation methods are based in grouping rectangular areas of cells of the finer resolution image to create a new map with larger cells. To do that, the function *aggregate()* of the package *raster* needs to know the factor of aggregation. In this case, the factor is 3 as it needs to go from 333m to 1km. In addition, *aggregate()* can perform the calculation using different functions. While the default is the average (*mean()*) it can work also with *modal()*, *max()* or *min()*.

As for the pixel coordinate definition for both the 300m and the 1km products, no proper aggregation of
the finer resolution product can be performed at the minimum and maximum latitude and longitude[^1]












[^1]: Erwin Wolters, Wouter Dierckx, Marian-Daniel Iordache, and Else Swinnen. (16/03/2018) PROBA-V Products User Manual v3.01. http://proba-v.vgt.vito.be/sites/proba-v.vgt.vito.be/files/products_user_manual.pdf
